---
output:
  github_document:
    html_preview: false
    md_extensions: -native_divs
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# jiebaRS

<div align="center">

<!-- badges: start -->
[![GitHub Stars](https://img.shields.io/github/stars/Yousa-Mirage/jiebaRS?style=social)](https://github.com/Yousa-Mirage/jiebaRS/stargazers)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Docs dev Version](https://img.shields.io/badge/docs-dev-blue.svg)](https://yousa-mirage.github.io/jiebaRS/)
[![Ask DeepWiki](https://deepwiki.com/badge.svg)](https://deepwiki.com/Yousa-Mirage/jiebaRS)
<!-- badges: end -->

[简体中文](./) | [English](README_en.md)

</div>

**jiebaRS** is a Rust-backed replacement of
[jiebaR](https://github.com/qinwf/jiebaR) for Chinese text segmentation,
part-of-speech tagging, and keyword extraction. It uses
[jieba-rs](https://github.com/messense/jieba-rs) as the segmentation engine,
bringing modern performance and maintainability.

[qinwf](https://github.com/qinwf) created
[jiebaR](https://github.com/qinwf/jiebaR) based on cppjieba many years ago,
which has long been the go-to tool for Chinese text segmentation in R. However,
[qinwf](https://github.com/qinwf) has stopped maintaining
[jiebaR](https://github.com/qinwf/jiebaR), and
[jiebaR](https://github.com/qinwf/jiebaR) has been removed from
[CRAN](https://cran.r-project.org/web/packages/jiebaR/index.html), making it
unavailable for direct installation. This has caused inconvenience and confusion
in R teaching and research. Therefore, based on the well-featured, performant,
and actively maintained Rust [jieba-rs](https://github.com/messense/jieba-rs)
crate, I developed **jiebaRS** to provide R users with a modern and easy-to-use
Chinese text segmentation tool, maintaining much compatibility with the original
[jiebaR](https://github.com/qinwf/jiebaR) API.

## Installation

### From CRAN

You can install the released version of jiebaRS from CRAN:

```r
install.packages("jiebaRS")
```

### From R-universe / R-multiverse

A prebuilt binary is hosted on R-universe and R-multiverse:

```r
install.packages("jiebaRS", repos = "https://yousa-mirage.r-universe.dev")
```

```r
install.packages("jiebaRS", repos = "https://community.r-multiverse.org")
```

### From Source Code

Install from source with [pak](https://pak.r-lib.org/) or
[remotes](https://remotes.r-lib.org/):

```r
pak::pak("Yousa-Mirage/jiebaRS")
# or
remotes::install_github("Yousa-Mirage/jiebaRS")
```

> **Note:** Building from source requires Rust tool chain to compile the Rust
> backend.

## Usage

```{r}
#| label: library

library(jiebaRS)
```

### Segmentation

Create a worker and segment a text:

```{r}
#| label: segment-basic

cutter <- worker()
segment("南京市长江大桥", cutter)
```

Batch segmentation supports multiple strings with `batch` aggregation. It will
automatically segment parallelly if more than one strings are provided, making
it much faster than jiebaR.

```{r}
#| label: segment-batch

texts <- c("南京市长江大桥。", "这是一个测试，小明很聪明。")

# list: one character vector per input string
segment_batch(texts, cutter, batch = "list")

# flatten: all tokens concatenated into one vector
segment_batch(texts, cutter, batch = "flatten")

# data.frame: doc_id + word columns
segment_batch(texts, cutter, batch = "data.frame")
```

If you want to segment a very long text parallelly, you can split it into 32\~64
chunks and then use `segment_batch()`.

### Speech Tagging

You can tag segmented words with part-of-speech (POS) tags using the `tagging()`
function:

```{r}
#| label: tagging

tagger <- worker(type = "tag")

# Default: named vector (terms as names, tags as values)
tagging("这是一个测试，小明很聪明。", tagger)

# data.frame: term + tag columns
tagging("这是一个测试，小明很聪明。", tagger, format = "data.frame")

# legacy: jiebaR-style (terms as values, tags as names)
tagging("这是一个测试，小明很聪明。", tagger, format = "legacy")
```

### Keyword Extraction

You can extract keywords using TF-IDF with the `keywords()` function:

```{r}
#| label: keywords

keys <- worker(type = "keywords", topn = 3)

text <- "今天纽约的天气真好啊，京华大酒店的张尧经理吃了一只北京烤鸭。后天纽约的天气不好，昨天纽约的天气也不好，北京烤鸭真好吃。"

# Named numeric vector (keyword -> weight)
keywords(text, keys)

# Data frame with term + weight columns
keywords_df(text, keys)
```

You can also use the TextRank algorithm with the `textrank()` function. This is
available in [Python's jieba](https://github.com/fxsjy/jieba) but not in jiebaR.

```{r}
#| label: textrank
#| 
ranker <- worker(type = "textrank", topn = 3)

textrank(text, ranker)

textrank_df(text, ranker)
```

### Custom Dictionaries

Load a custom **main dictionary** (`dict` --- replaces the embedded dictionary)
or a **user dictionary** (`user` --- appends to the main dictionary). Both files
use the line format: `word [freq] [tag]`.

```{r}
#| label: custom-dict

# User dictionary: add new words to the default dictionary
user_file <- withr::local_tempfile()
writeLines(c("量子机器狗 1000 n", "超导量子比特 1000"), user_file, useBytes = TRUE)

cutter2 <- worker(user = user_file)
segment("量子机器狗和超导量子比特", cutter2)
```

Add words dynamically with `new_user_word()` (alias: `add_word()`):

```{r}
#| label: new-user-word

cutter3 <- worker()
segment("量子机器狗和超导量子比特", cutter3)

new_user_word(cutter3, "量子机器狗", "n")
add_word(cutter3, "超导量子比特", "n")  # alias
segment("量子机器狗和超导量子比特", cutter3)
```

### Stop Words

Supply stop words as a character vector via the `stop_word` parameter or a file
path via the `stop_word_file` parameter. Stop words are filtered from
segmentation and keyword extraction results.

```{r}
#| label: stop-words

cutter4 <- worker(stop_word = c("这是", "一个"))
segment("这是一个测试", cutter4)
```

### Custom IDF + `get_idf()`

Compute an IDF dictionary from your own corpus of segmented documents, then use
it for TF-IDF keyword extraction:

```{r}
#| label: custom-idf

docs <- list(
  c("北京", "烤鸭", "纽约"),
  c("北京", "天气"),
  c("纽约", "天气")
)

idf_file <- withr::local_tempfile()
get_idf(docs, path = idf_file)

keys_custom <- worker(type = "keywords", topn = 3, idf = idf_file)
keywords(text, keys_custom)
```

### Word Frequency & N-grams

Two small but useful functions are provided for word frequency and n-gram
counts:

```{r}
#| label: freq-ngrams

tokens <- segment("南京市长江大桥南京市", worker())

# Word frequency
freq(tokens)

# Sorted by descending frequency
freq(tokens, sort = TRUE)

# N-gram counts (default: bigrams)
count_ngrams(tokens, n = 2)

# Multiple n sizes, as a named vector
count_ngrams(tokens, n = 1:2, format = "vector")
```

## Compare with jiebaR

jiebaRS and jiebaR are both based on the jieba segmentation algorithm, but there
are inevitably subtle differences between the Rust backend (jieba-rs) and the
C++ backend (cppjieba). The results below are measured on the full text of
*Fortress Besieged* (围城) and *Dream of the Red Chamber* (红楼梦).

### Segmentation Similarity

| Corpus | Characters | jiebaRS tokens | jiebaR tokens | jiebaRS vocab | jiebaR vocab | Vocab Jaccard |
|--------|----------:|--------------:|-------------:|--------------:|-------------:|:-------------:|
| Fortress Besieged   | 246,871 | 128,985 | 129,560 | 18,375 | 18,794 | 0.929 |
| Dream of the Red Chamber | 860,933 | 451,792 | 451,228 | 44,634 | 45,596 | 0.865 |

The total token counts are close (within 0.2%), and the vocabularies overlap
heavily. The main differences come from the different segmentation granularity
of some proper nouns (e.g. jiebaRS merges the protagonist name "鸿渐" into a
single word, while jiebaR splits it into "鸿" + "渐"), as well as minor
differences in HMM-based OOV (out-of-vocabulary) boundary inference.

### Segmentation Speed

| Corpus | Input mode | jiebaRS (s) | jiebaR (s) | Speedup |
|--------|-----------|------------:|-----------:|:-------:|
| Fortress Besieged | whole text | 0.039 | 0.065 | 1.66x |
| Fortress Besieged | 100k sentences (parallel) | 0.246 | 2.930 | 11.91x |
| Dream of the Red Chamber | whole text | 0.129 | 0.246 | 1.91x |
| Dream of the Red Chamber | 100k sentences (parallel) | 0.639 | 4.601 | 7.20x |

For a single long text, jiebaRS is about 1.7\~1.9x faster; when the input is many
short sentences that are segmented in parallel, jiebaRS can reach 7\~12x speedup.

## Acknowledgments

jiebaRS builds on the work of open-source projects:

- **[jieba-rs](https://github.com/messense/jieba-rs)**: messense and
  contributors for the Rust port of the Jieba engine.
- **[jiebaR](https://github.com/qinwf/jiebaR)**: qinwf for the original R
  package that jiebaRS replaces.
- **[extendr](https://github.com/extendr/extendr)**: the extendr team for making
  Rust--R interoperability practical.
- The broader **Rust** and **R** communities.

Without these projects, jiebaRS would not exist.

## License

jiebaRS is licensed under the [MIT License](LICENSE.md).
